---
title: "Homework 1"
author: "Irene Wen"
date: "2026-09-18"
output: github_document
---

```{r load_libraries, message = FALSE}
library(tidyverse)
```

# Problem 1
```{r load_penguins}
data("penguins", package = "palmerpenguins")
```
The `penguins` dataset contains measurements of penguins observed in the
Palmer Archipelago, Antarctica. It has `r nrow(penguins)` observations and
`r ncol(penguins)` variables.

The dataset includes penguins from three species
(`r paste(levels(pull(penguins, species)), collapse = ", ")`) living on
three islands (`r paste(levels(pull(penguins, island)), collapse = ", ")`).
Important variables include bill length (`bill_length_mm`), bill depth
(`bill_depth_mm`), flipper length (`flipper_length_mm`), body mass
(`body_mass_g`), `sex`, and `year`. Observations were collected from
`r min(pull(penguins, year))` to `r max(pull(penguins, year))`.

The mean flipper length is
`r round(mean(pull(penguins, flipper_length_mm), na.rm = TRUE), 2)` mm.

```{r penguins_scatterplot}
penguins_scatterplot =
  ggplot(
    penguins,
    aes(
      x = bill_length_mm,
      y = flipper_length_mm,
      color = species
    )
  ) +
  geom_point()

penguins_scatterplot

ggsave(
  "penguins_scatterplot.png",
  plot = penguins_scatterplot,
  width = 7,
  height = 5
)
```

The scatterplot shows a generally positive relationship between bill length and flipper length, although the three penguin species form distinct clusters. Adelie penguins tend to have shorter bills and flippers, Gentoo penguins generally have the longest flippers, and Chinstrap penguins tend to have relatively long bills but intermediate flipper lengths. The warning indicates that observations with missing values in bill_length_mm or flipper_length_mm were removed from the plot.

# Problem 2
```{r create_problem2_df}
set.seed(1)

problem2_df =
  tibble(
    norm_samp = rnorm(10),
    norm_samp_pos = norm_samp > 0,
    vec_char = c("a", "b", "c", "d", "e", "f", "g", "h", "i", "j"),
    vec_factor = factor(
      c("low", "medium", "high", "low", "medium",
        "high", "low", "medium", "high", "low")
    )
  )

problem2_df
```

```{r take_means}
mean(pull(problem2_df, norm_samp))
mean(pull(problem2_df, norm_samp_pos))
mean(pull(problem2_df, vec_char))
mean(pull(problem2_df, vec_factor))
```
Taking the mean works for the numeric and logical variables but not for the
character and factor variables. The mean of `norm_samp` is
`r round(mean(pull(problem2_df, norm_samp)), 2)`, which is close to 0, as
expected for a sample from a standard Normal distribution. The mean of
`norm_samp_pos` is `r mean(pull(problem2_df, norm_samp_pos))`; because R
treats `TRUE` as 1 and `FALSE` as 0, this is the proportion of sampled
values that are greater than 0.

For `vec_char` and `vec_factor`, `mean()` returns `NA` with a warning that
the argument is not numeric or logical. Character values have no numeric
meaning, and a factor represents categories rather than quantities, so an
arithmetic mean is not defined for either.


```{r coerce_to_numeric, results = "hide", warning = FALSE}
as.numeric(pull(problem2_df, norm_samp_pos))
as.numeric(pull(problem2_df, vec_char))
as.numeric(pull(problem2_df, vec_factor))
```

Applying `as.numeric()` converts the logical variable to 1 for `TRUE` and 0
for `FALSE`. The sample contains
`r sum(pull(problem2_df, norm_samp_pos))` ones out of
`r nrow(problem2_df)` values, which matches the mean of `norm_samp_pos`
computed above. This explains why taking the mean of a logical variable
works: it gives the proportion of `TRUE` values.

The character variable cannot be converted, so every value becomes `NA`
(with a warning that NAs were introduced by coercion). This is consistent
with `mean()` returning `NA` for the character variable, since there are no
numeric values to average.

The factor variable is converted to integer codes. R stores factor levels in
alphabetical order (`r paste(levels(pull(problem2_df, vec_factor)), collapse = ", ")`),
so "high" becomes 1, "low" becomes 2, and "medium" becomes 3. These codes
are only labels used internally and do not even follow the natural ordering
of low < medium < high. `mean()` does not convert factors automatically,
which is why it returned `NA`, and even after explicit conversion, averaging
these codes would not produce a meaningful quantity.