---
title: "Homework 1"
author: "Irene Wen"
date: "2026-09-18"
output: github_document
---

```{r load_libraries, message = FALSE}
library(tidyverse)
```

# Problem 1
```{r load_penguins}
data("penguins", package = "palmerpenguins")
levels(pull(penguins, species))
levels(pull(penguins, island))

```
The `penguins` dataset includes the species, island, bill length, bill depth,
flipper length, body mass, sex, and year for each penguin. There are three
species ("Adelie"    "Chinstrap" "Gentoo" ) and three islands
("Biscoe"    "Dream"     "Torgersen").

```{r}
nrow(penguins)
ncol(penguins)
penguins |> pull(flipper_length_mm) |> mean(na.rm = TRUE) |> round(2)
```
The dataset has 344 rows and 8 columns, so
there are 344 penguins and 8 variables.

The mean flipper length is 200.92 mm.


```{r penguins_scatterplot}
penguins_scatterplot =
  ggplot(
    penguins,
    aes(
      x = bill_length_mm,
      y = flipper_length_mm,
      color = species
    )
  ) +
  geom_point()

penguins_scatterplot

ggsave(
  "penguins_scatterplot.png",
  plot = penguins_scatterplot,
  width = 7,
  height = 5
)
```


# Problem 2
```{r create_problem2_df}
set.seed(1)

problem2_df =
  tibble(
    norm_samp = rnorm(10),
    norm_samp_pos = norm_samp > 0,
    vec_char = c("a", "b", "c", "d", "e", "f", "g", "h", "i", "j"),
    vec_factor = factor(
      c("low", "medium", "high", "low", "medium",
        "high", "low", "medium", "high", "low")
    )
  )

problem2_df
```

```{r take_means}
mean(pull(problem2_df, norm_samp))
mean(pull(problem2_df, norm_samp_pos))
mean(pull(problem2_df, vec_char))
mean(pull(problem2_df, vec_factor))
```
Taking the mean works for the numeric and logical variables. For the logical
variable, the mean is the proportion of `TRUE` values. It doesn't work for the
character and factor variables, which return `NA` with a warning that the
argument is not numeric or logical.


```{r, eval = FALSE}
as.numeric(pull(problem2_df, norm_samp_pos))
as.numeric(pull(problem2_df, vec_char))
as.numeric(pull(problem2_df, vec_factor))
```
The logical variable becomes 1 for `TRUE` and 0 for `FALSE`. The character
variable becomes all `NA`, because letters can't be converted to numbers. The
factor variable becomes the numbers 1, 2, and 3, which are the codes R uses for
its levels (high, low, medium in alphabetical order).

This helps explain the means. The logical variable works because it is stored
as 0s and 1s. The character variable has no numeric values, so there is nothing
to average. The factor does have numeric codes underneath, but `mean()` doesn't
convert factors, and the codes only label categories, so their mean wouldn't
mean anything.